Illustrating Model Properties

### 2024-06-07 manuscript.qmd
simpler legend in risk prediction

![Histograms over the predicted risk of contralateral LNL II invovlement, shown for all eight possible combinations of the risk factors T-category (early: top row, advanced: bottom row), midline extension (absent: left column, present: right column), and ipsilateral diagnosis (ipsilateral N0: blue, ipsilateral at least LNL III diagnosed: orange). The provided diagnosis always included a clinically node negative contralateral neck. Each of the three risk factors significantly increase the predicted risk for occult disease](figures/fig-model-risks.png){#fig-model-risks}

>>>>

![Histograms over the predicted risk of occult contralateral LNL II invovlement, shown for some combinations of T-category, tumor lateralization, and ipsilateral clinical involvement. The contralateral side was always assumed to be clinically negative.](figures/fig-model-risks.png){#fig-model-risks width=50%}

### 2024-06-07 model.qmd
simpler legend in risk prediction

>>>>

# Risk Predictions

Now, we want to make actual use of the model and predict the risk for occult diease given some diagnoses.

```{python}
import numpy as np
import matplotlib.pyplot as plt
from matplotlib.legend_handler import HandlerTuple
from lyscripts.scenario import Scenario
from lyscripts.plot.utils import Histogram, draw, COLORS, split_legends
from scripts import paths, shared

all_false = {lnl: False for lnl in ["I", "II", "III", "IV", "V"]}

plt.rcParams.update(shared.get_fontsizes())
plt.rcParams.update(shared.get_figsizes(width=17./2., aspect_ratio=2.))

fig, ax = plt.subplots()
ax.set_xlim(0., 15.)

contents = []
defaults = {
  "density": True,
  "histtype": "stepfilled",
  "alpha": 0.7,
  "bins": 50,
}

scenario = Scenario(
  t_stages=["early"],
  midext=False,
  diagnosis={
    "ipsi": {"CT": all_false},
    "contra": {"CT": all_false},
  },
  involvement={"contra": {"II": True}},
)
contents.append(Histogram.from_hdf5(
  filename=paths.model_dir / "full" / "risks.hdf5",
  dataname=scenario.md5_hash("risks"),
  color=COLORS["green"],
  label="early, lateralized; N0",
))

scenario = Scenario(
  t_stages=["late"],
  midext=False,
  diagnosis={
    "ipsi": {"CT": {"III": True}},
    "contra": {"CT": all_false},
  },
  involvement={"contra": {"II": True}},
)
contents.append(Histogram.from_hdf5(
  filename=paths.model_dir / "full" / "risks.hdf5",
  dataname=scenario.md5_hash("risks"),
  color=COLORS["blue"],
  label="advanced, lateralized; ipsi LNL III",
))

scenario = Scenario(
  t_stages=["early"],
  midext=True,
  diagnosis={
    "ipsi": {"CT": all_false},
    "contra": {"CT": all_false},
  },
  involvement={"contra": {"II": True}},
)
contents.append(Histogram.from_hdf5(
  filename=paths.model_dir / "full" / "risks.hdf5",
  dataname=scenario.md5_hash("risks"),
  color=COLORS["orange"],
  label="early, midline extension; N0",
))

scenario = Scenario(
  t_stages=["late"],
  midext=True,
  diagnosis={
    "ipsi": {"CT": {"III": True}},
    "contra": {"CT": all_false},
  },
  involvement={"contra": {"II": True}},
)
contents.append(Histogram.from_hdf5(
  filename=paths.model_dir / "full" / "risks.hdf5",
  dataname=scenario.md5_hash("risks"),
  color=COLORS["red"],
  label="advanced, midline extension; ipsi LNL III",
))

draw(ax, contents, xlims=(0., 15.))
ax.legend(labelspacing=0.3)
ax.set_xlabel("risk [%]")
ax.set_yticks([])

plt.savefig(paths.figure_dir / "fig-model-risks.png", dpi=400)
plt.show()
```
